{CPU-Scheduling}

### 2024-12-14 LaTeX/Sources/Advanced.tex
started lottery sched

Tickets are a form of currency for the computer.
These can be handed out by the scheduler and give a process the right to use a resource.
Usually this is done proportianally to the amount of tickets.
Meaning a higher share of tickets will lead to a higher amount of CPU time.
Logically the response time is reverse proportianal.
These tickets enable easy comparison between processes and their resources.
One important difference between a currency in real life and tickets is that tickets are consumed when used.
Practically this means that if you buy something you won't lose your bill.
This property leads to the tickets representing the share of the CPU that the job has.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{Assets/Ticket-Currency.pdf}
    \caption{Example of a Custom Ticket Currency}
    \label{fig:ticket-currencies}
\end{figure}

Another benefit comes with the ticket inflation or deflation operation, which can only be used by the owner of the currency.
This is the case, because it involves creating or destroying tickets. 
Still unlike the ticket transfer, ticket inflation does not need a sender and reciever.
The selected process just get more tickets assigned.
With this the parent process changes the ration of ownership and therefore boosts the selected job.

>>>>

Tickets are a form of currency for the computer.
These can be handed out by the scheduler and give a process the right to use a resource.
Usually this is done proportianally to the amount of tickets.
Meaning a higher share of tickets will lead to a higher amount of CPU time.
Logically the response time is reverse proportional.
These tickets enable easy comparison between processes and their resources.
One important difference between a currency in real life and tickets is that tickets are consumed when used.
Practically this means that if you buy something you won't lose your bill.
This property leads to the tickets representing the share of the CPU that the job has.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{Assets/Ticket-Currency.pdf}
    \caption{Example of a Custom Ticket Currency}
    \label{fig:ticket-currencies}
\end{figure}

Another benefit comes with custom currencies: the ticket inflation or deflation operation.
One needs a custom currency, because it involves creating or destroying tickets. 
Still unlike the ticket transfer, ticket inflation does not need a sender and reciever.
The selected process just get more tickets assigned.
With this the parent process changes the ration of ownership and therefore boosts the selected job.

\section{Lottery Scheduling}

Like other things in computer science the name already says a lot. 
Here randomization is exploited to achieve the proportianal share in runtime.
The concept that if even over a small scale there will be unavoidable errors, the longer the scheduler runs the more accurate the share becomes.
Think of the tickets as lottery tickets.
If the randomization is fair, than the more tickets you have the more you will win.
There is a downside though.
The randomizer algorithm can get pretty expensive.
There is a way to reuse one random seed though, but we will only quickly touch upon it later.

For an example implementation in C take a look at figure \ref{code:lottery-sched}
This is taken from phd-mit-tr677.pdf (CITATION)

\begin{figure}[h]
    \begin{minted}[mathescape,
        linenos,
        numbersep=5pt,
        gobble=2,
        frame=lines,
        framesep=2mm,
        ]{cpp}
  /* Implementation of List-Based Lottery Scheduling Algorithm (Written by: pdh-mit-tr667.pdf CITATION) */

  /*per-client state */
  typedef struct {
    ...
    int tickets;
  } *client_t;

  /* current resource owner */
  client_t current;

  /* list of clients competing for resource */
  list_t list;

  /*global tick et sum */
  int global_tickets= 0;

  /* initialize client withspecified allocation */
  void client_init(client_t c, int tickets)
  {
    /*initialize client state, updateglobal sum */
    c->tickets = tickets;
    global_tickets += tickets;
    /*join competition for resource */
    list_insert(list, c);
  }

  void allocate()
  {
    int winner, sum;
    client_t c;
    /* randomly select winning tick et */
    winner = fast_random() % global_tickets;
    /* searchlist to find client with winning tick et */
    sum = 0;
    for (c = list_first(list);
         c != NULL;
         c = list_next(list, c))
    {
        /*update running sum, stop at winner */
        sum += c->tickets;
        if (sum > winner)
            break;
    }
    /*grant resource to winner forquantum */
    current = c;
    use_resource(current);
  }
    \end{minted}
    \label{code:lottery-sched}
    \caption{C: Lottery Scheduling}
\end{figure}

One could also use Binary trees...
